---
title: "Shared Actor Network"
author: "Gordon Blasco"
date: "5/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Making a network graph of popular high rated movies from IMBD connected by their shared "principle" directors, producers, writers, and actors. All data from IMBD's data portal: https://www.imdb.com/interfaces/

### Libraries needed
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidygraph)
library(networkD3)
library(beepr)
```

### Read in data (I could not upload these to github due to their size, so if you want to make this you will need to download the 4 data files independently)
```{r eval=FALSE}
# I added the name of the dataset as listed and availible for download at: https://www.imdb.com/interfaces/ 
ratings <- read_tsv("ratings.tsv") %>% # title.ratings.tsv.gz (movie ratings)
  rename(titleId = tconst)

names <- read_tsv("name_basics.tsv") %>% # name.basics.tsv.gz (actor names)
  select(nconst, primaryName)

title_basics <- read_tsv("title_basics.tsv") %>% # title.basics.tsv.gz (movie titles & information)
  rename(titleId = tconst)

principals <- read_tsv("title_principals.tsv") %>% # title.principals.tsv.gz (actor, writer, director codes)
  rename(titleId = tconst)

beep(8) # These datasets are large and took my computer a while to read in. So, I set a beeper to go off once they are done so I could prep dinner without checking on my computer (I made wontons!)
```

### Combine datasets
```{r eval=FALSE}
films <- title_basics %>% 
  filter(titleType == "movie") %>% # Remove TV shows and shorts
  filter(startYear <= 2018) %>% # Keep movies made before 2018
  inner_join(., ratings, by = "titleId")
  

films <- inner_join(principals, movies, key = "titleID") %>% 
  inner_join(., names)

```


### Selecting films
I'm interested in movies that have a rating of 8 stars or higher, were made after 2000, and that are popular enough to have at least 500 votes. 
```{r eval=FALSE}
selected_films <- films %>% 
  filter(averageRating >= 8 & numVotes >= 500 & startYear >= 2000) %>% 
  select(primaryTitle, primaryName) # Only the name of the movie and the actor will be needed going forward
```

```{r include=FALSE}
selected_films <- read_csv("films_small.csv")
```


### Link films with 
```{r}
films1 <- selected_films # Create two copys of the dataset to combine by matching actor, writer, director
films2 <- selected_films %>% 
  rename(title = primaryTitle) 

simple_films <- dplyr::full_join(films1, films2, by = "primaryName") 

tidy_fims <- simple_films %>% 
  filter(!primaryTitle == title) %>% # Remove films connected to themselves
  distinct(primaryTitle, title) # Only keep unique connections, no need to have multiples of the same connection. 
```


```{r}
simpleNetwork(tidy_fims, zoom = TRUE)
```

